---
title: "Multiple Indicator Sigma Chart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multiple Indicator Sigma Chart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = TRUE}
library(qiverse.qimisc)
```

This package provides functions to prepare data ready for the Multiple Indicator Sigma Chart (MISC), and chart this using plotly.

## Prepare MISC Data

The data that feeds into the Multiple Indicator Sigma Chart requires preparation using the `misc_prep_data` function. This takes two inputs:

-   `funnel_data` - A data.frame/data.table of the funnel data with the columns indicator, group, numerator and denominator.

-   `indicator_data` - A data.frame/data.table of the indicator metadata with the columns indicator_theme, indicator, multiplier, data_type and betteris.

Examples of these datasets are below:

```{r}
example_funnel_data
```

```{r}
example_indicator_data
```

These datasets must be in the above format to feed into the MISC data preparation.

```{r}
misc_data <- misc_prep_data(example_funnel_data, example_indicator_data)
misc_data
```

## Charting MISC in Plotly

This package provides a function to plot the MISC output using Plotly. An example of this is as follows:

```{r, fig.height = 4, fig.width = 10, fig.align = "center"}
misc_plotly(misc_data[group == "A"])
```
