---
title: "Quality Improvement Pattern Detection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quality Improvement Pattern Detection}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = TRUE}
library(qiverse.qipatterns)
library(qiverse.data)
```

This package provides functions to detect Funnel plot and Statistical
Process Control patterns.

# Most Recent Pattern Detection for SPC and Funnels

The `pattern_detection` function extracts the most recent detected SPC
and Funnel pattern for each indicator and group combination.

## Prepare Data

The data that feeds into the `pattern_detection` function should be in a
long format and contain the following columns:

-   `indicator` - A vector to denote the indicator group. Used to
    calculate aggregate SPC and Funnel Plots
-   `group` - A vector of group names. Used to aggregate and group
    points on plots
-   `period_end` - A date (or character of format "yyyy-mm-dd") for the
    end date of the period.
-   `numerator` - A vector of the numerator. Used to calculate the value
    for SPC and Funnel plots
-   `denominator` - A vector of the denominator Used to calculate the
    value for SPC and Funnel plots
-   `multiplier` - Scale relative risk and funnel by this factor.
-   `better_is` - A string identifying the direction that is favourable
    for the indicator. "Higher" for points below the lower control limit
    to be unfavourable, "Lower" for points above the upper control limit
    to be unfavourable, and "Neutral" if the direction is not stated.
-   `parent_group_name` - A vector of parent group names which are to be
    displayed in the tooltip. These are the major categories for the
    groups.
-   `spc_chart_type` - A string identifying the type of SPC chart.
-   `funnel_chart_type` - A string identifying the type of funnel plot.
-   `indicator_group` - A vector used for indicator grouping. i.e.
    grouping together maternity related indicators into a Maternity
    group.
-   `indicator_name` - A vector of descriptive names for indicator.
-   `group_name` - A vector of descriptive names for groups Default
    group variable
-   `funnel_data_points` - A vector of "Yes" or NA, marks which data
    points are included in the funnel plot calculation.

An example of this is shown below using the `example_spc_data` dataset:

```{r}
head(example_spc_data)
```

## Pattern Detection Function

The dataset is then passed to the `pattern_detection` function to detect
SPC and Funnel patterns:

```{r pattern-detection, include = TRUE}
pattern_detection_data <- pattern_detection(
  indicator = example_spc_data$indicator,
  group = example_spc_data$group,
  period_end = example_spc_data$period_end,
  numerator = example_spc_data$numerator,
  denominator = example_spc_data$denominator,
  multiplier = example_spc_data$multiplier,
  better_is = example_spc_data$better_is,
  spc_chart_type = example_spc_data$spc_chart_type,
  funnel_chart_type = example_spc_data$funnel_chart_type,
  indicator_group = "All", # Allocates indicators into one parent group named All
  indicator_name = example_spc_data$indicator,
  group_name = example_spc_data$group,
  funnel_data_points = "Yes", # Includes all data points in the funnel plot
  trend_size = 5,
  shift_size = 7
)
head(pattern_detection_data, 20)
```

The above code will also work if you are not interested in generating
funnel data points, and only have data for one group.

```{r pattern-detection-groupA, include = TRUE}
pattern_detection_data_groupA <- pattern_detection(
  indicator = example_spc_data[group == "A"]$indicator,
  group = example_spc_data[group == "A"]$group,
  period_end = example_spc_data[group == "A"]$period_end,
  numerator = example_spc_data[group == "A"]$numerator,
  denominator = example_spc_data[group == "A"]$denominator,
  multiplier = example_spc_data[group == "A"]$multiplier,
  better_is = example_spc_data[group == "A"]$better_is,
  spc_chart_type = example_spc_data[group == "A"]$spc_chart_type,
  funnel_chart_type = example_spc_data[group == "A"]$funnel_chart_type,
  indicator_group = "All", # Allocates indicators into one parent group named All
  indicator_name = example_spc_data[group == "A"]$indicator,
  group_name = example_spc_data[group == "A"]$group,
  funnel_data_points = "Yes", # Includes all data points in the funnel plot
  trend_size = 5,
  shift_size = 7
)
head(pattern_detection_data_groupA, 20)
```

As displayed above, the output is one record for each indicator and
group combination. The columns show:

-   The sum of the numerator

-   The sum of the denominator

-   The most recent astronomical point period end date

-   The most recent shift period end date

-   The most recent trend period end date

-   The most recent two-in-three period end date

-   The most recent funnel plot outlier period end date

# All SPC Pattern Detection

The `pattern_rules` function extracts all patterns relating to
Statistical Process Control patterns for Astronomical Points, Trends,
Shifts and Two-in-three.

## Pattern Rules Function

The `pattern_rules` function uses data.table to parallelise the SPC
limit calculations and pattern detections. This requires a unique.key to
be created to identify which records are related to each other, and
should be considered in the same SPC calculations. An example is as
follows:

```{r pattern_rules, include = TRUE}
pattern_rules_data <- qiverse.qipatterns::pattern_rules(
  numerator = example_spc_data$numerator,
  denominator = example_spc_data$denominator,
  period_end = as.character(example_spc_data$period_end),
  unique_key = example_spc_data[, paste0(indicator, "_", group)],
  spc_chart_type = example_spc_data$spc_chart_type,
  multiplier = example_spc_data$multiplier,
  better_is = example_spc_data$better_is,
  trend_size = 5,
  shift_size = 7
)
head(pattern_rules_data, 72)
```

Similarly, we can generate SPC patterns and limit calculations for a
single indicator and group combination.

```{r pattern_rules_SMR_A, include = TRUE}
example_spc_data_SMR_A <- example_spc_data[indicator == "Standardised Mortality Ratio" & group == "A"]
pattern_rules_data_SMR_A <- qiverse.qipatterns::pattern_rules(
  numerator = example_spc_data_SMR_A$numerator,
  denominator = example_spc_data_SMR_A$denominator,
  period_end = as.character(example_spc_data_SMR_A$period_end),
  unique_key = NA,
  spc_chart_type = "i",
  multiplier = 100,
  better_is = "Lower",
  trend_size = 5,
  shift_size = 7
)
head(pattern_rules_data_SMR_A, 36)
```
