---
title: "Downloading and Uploading to Sharepoint"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Downloading and Uploading to Sharepoint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(qiverse.azure)
library(qiverse.sharepoint)
```

The series of functions in `qiverse.sharepoint` are wrappers of `Microsoft365R` functions to perform some key tasks such as downloading and uploading files to SharePoint, and downloading data from a SharePointlist. The `Microsoft365R` package provides a lot of functionality that leverages off the Microsoft Graph API, including the ability to create, modify and delete lists, manage SharePoint sites and many more.

This package provides three SharePointfunctions:

-   download_sharepoint_file

-   upload_sharepoint_file

-   download_sharepoint_list

Each of these will be covered in more detail below.

## Download SharePoint File

This function uses the `Microsoft365R` package to download any file in SharePoint, which can either be loaded directly in-memory to R or to a specific location on your local computer. The `download_sharepoint_file` function looks as follows:

``` r
download_sharepoint_file(
  site_url,
  file_url,
  token,
  download,
  download_dest
)
```

With descriptions for the arguments as follows:

-   `site_url` - This is the URL when first accessing the SharePoint team via your web browser (Home page)

-   `file_url` - This is a direct link to the file to be downloaded. It can be obtained by navigating to the file you want to download, right-clicking and selecting "Copy Link"

-   `token` - The token generated with the correct SharePoint permissions. Use get_az_tk('sp") to create this token.

-   `download` - Set this to FALSE to download the file into a temporary file. This temporary file can then be read in using any standard data reading function. Set this to TRUE to download the file to your computer.

-   `download_dest` - This option is only active when download is set to TRUE. When download_location is set to NULL, the file is just downloaded into the current working directory. Otherwise, specify the destination path to save the file.

The three use cases for this function is as below:

### Download SharePoint File to Working Directory

``` r
# Create SharePoint azure token
library(qiverse.azure)
tk <- get_az_tk('sp')

# Download file to working directory from SharePoint site
library(qiverse.sharepoint)
download_sharepoint_file(
  site_url = "https://myexample.sharepoint.com/sites/Example-Files/"),
  file_url = paste0("https://myexample.sharepoint.com/:x:/r/sites/Example-Files/Shared%20Documents/Example_datasets/example_data.csv?d=XXXXXXXXXXXXXXXXXXXXXXXXXXXX",
  token = tk,
  download = TRUE
)
```

### Download SharePoint File to Selected Directory

``` r
# Create SharePoint azure token
library(qiverse.azure)
tk <- get_az_tk('sp')

# Download file to C:/ drive from SharePoint
library(qiverse.sharepoint)
download_sharepoint_file(
  site_url = "https://myexample.sharepoint.com/sites/Example-Files/"),
  file_url = paste0("https://myexample.sharepoint.com/:x:/r/sites/Example-Files/Shared%20Documents/Example_datasets/example_data.csv?d=XXXXXXXXXXXXXXXXXXXXXXXXXXXX",
  token = tk,
  download = TRUE,
  download_dest = 'C:/cabg.csv'
)
```

### Download SharePoint File and Load In-Memory to R

``` r
# Create SharePoint azure token
library(qiverse.azure)
tk <- get_az_tk('sp')

# Example to download to tempfile, to load straight into R memory
library(qiverse.sharepoint)
data <- download_sharepoint_file(
  site_url = "https://myexample.sharepoint.com/sites/Example-Files/"),
  file_url = paste0("https://myexample.sharepoint.com/:x:/r/sites/Example-Files/Shared%20Documents/Example_datasets/example_data.csv?d=XXXXXXXXXXXXXXXXXXXXXXXXXXXX",
  token = tk,
  download = FALSE
) |>
  # Pipe in the approprate function to read in the data
  read.csv()
```

## Upload SharePoint File

This function uses the `Microsoft365R` package to upload any file into SharePoint. The `upload_sharepoint_file` function looks as follows:

``` r
upload_sharepoint_file(
  src,
  site_url,
  dest_fldr_url,
  token
)
```

With descriptions for the arguments as follows:

-   `src` -The file path to the local file that is being uploaded to SharePoint

-   `site_url` - This is the url when first accessing the SharePoint team via your web browser (Home page)

-   `file_url` - This is a direct link to the SharePoint folder where the file will be uploaded to. It can be obtained by navigating to the folder you want to upload to, right-clicking and selecting "Copy Link"

-   `token` - The token generated with the correct SharePoint permissions. Use get_az_tk('sp") to create this token.

Here is an example code block of this function:

``` r
# Create SharePoint azure token
library(qiverse.azure)
tk <- get_az_tk('sp')

# Example to upload local file to SharePoint folder
library(qiverse.sharepoint)
upload_sharepoint_file(
  src = 'TestFiles/test.txt',
  site_url = "https://myexample.sharepoint.com/sites/Example-Files/"),
  file_url = paste0("https://myexample.sharepoint.com/:x:/r/sites/Example-Files/Shared%20Documents/Example_datasets/example_data_amended.csv?d=XXXXXXXXXXXXXXXXXXXXXXXXXXXX",
  token = tk
)
```

## Download SharePoint List

The `download_sharepoint_list` function is able to download SharePoint lists into a data.frame object. The function looks as follows:

``` r
download_sharepoint_list(
  list_url,
  token
)
```

With the arguments defined as:

-   `list_url` - The URL of the list when accessing the list via your web browser.

-   `token` - The token generated with the correct SharePoint permissions. Use get_az_tk('sp") to create this token.

Here is an example code block of this function:

``` r
# Get the token for SharePoint
library(qiverse.azure)
tk <- get_az_tk('sp')

# Download a particular view for the SharePoint list
library(qiverse.sharepoint)
list_data <- download_sharepoint_list(
  paste0("https://myexample.sharepoint.com/sites/",
  "Indicator%20Metadata/Lists/Indicators/",
  "Latest.aspx"),
  token = tk
)

# Download the default view for the SharePoint list
list_data <- download_sharepoint_list(
  paste0("https://myexample.sharepoint.com/sites/",
  "Indicator%20Metadata/Lists/Indicators"),
  token = tk
)
```

## Other SharePoint Functionality

The functions available in `qiverse.sharepoint` serve as easy-to-use wrappers of the `Microsoft365R` package, to conduct key functions such as downloading and uploading SharePoint files, and downloading SharePoint lists. More functionality is available directly in the package such as create, modify and delete lists, and managing SharePoint sites.
